Composing like a human

### 2019-05-16 paper/main.tex
thesis == done (i hope)

\usepackage[colorlinks=true, citecolor=base08, linkcolor=base08, urlcolor=base08]{hyperref}

% Essentials
\usepackage{tabularx}
\usepackage{supertabular}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{lscape}
\usepackage[ruled]{algorithm2e}

@@

@@

%TC:ignore
\chapter*{Abstract}
Experiments in Deep Learning are marked by scant explorations outside of the classification task, and even more so when trying to overcome some of the limitations imposed by gradient-based optimization algorithms (such as the uncertainty of convergence when dealing with non-convexity). With the resurgence of generative models and recent advancements in meta-learning we are faced two exciting ares of research, the intersection of which could be a promising area when investing general-purpose A.I. systems. Inspired by \textcite{clouatre_figr_2019} and \textcite{zhang_metagan_2018}, this paper proposed a new line of inquiry where these two fields could be useful specifically, in the musical domain. As such, the paper sought to establish the extent to which meta-learned generative models are comparable to a model trained under normal circumstances, and to real musical performances.

@@

When compared to the baseline, the two Reptile-based models have show to be quite competitive however, they failed to attain the same level of mastery when compared to the test samples. However, it is hypothesized that more time involved into training them could narrow the gap between the models and the real musical performances. Thus, the results of these experiments solidify the intuition that meta-learning algorithms are a viable alternative to traditional training methods, in the context of generative models.

@@

\chapter{A hazy shade of winter} \label{chap:intro}
%Simon and Garfunkel - A hazy shade of winter

@@

Broadly speaking, DL as a field, promises to solve tasks which are easily (intuitively) performed by people, but hard to formalize \parencite[e.g., recognizing faces or spoken word;][]{goodfellow_deep_2016}. The solution comprises of enabling computers to learn from experience and understand the world through the discovery of hierarchical relationships between concepts \parencite{lecun_deep_2015}. This way, there is no need for manual input on the sort of knowledge that the computer needs \parencite{goodfellow_deep_2016}. A more simplistic and concise interpretation of DL would be to see it as a statistical technique for identifying patterns in sample data using large (deep) ANNs \parencite{marcus_deep_2018}.

@@

\section{Strange Fruit}\label{sec:no_disc}
% Billie Holiday - Strange Fruit

@@

Regardless, successes in generative models have been either scarce or unattractive, with Fully Visible Belief Networks \parencite{frey_does_1996} being the most suitable (and popular) method for the longest run. Given the age and - alleged - popularity, a valid question to ask is why things haven't moved that much since then. As it turn out, this method requires generating one sample at a time (ramping up the cost of the process to $O(n)$). Coupled with the fact that the computation is done through a neural network, creating $n$ samples requires a substantial amount of time \parencite{goodfellow_nips_2016}. This discussion is not exhaustive; for a comprehensive list of various generative approaches and their downsides, refer to \textcite{goodfellow_nips_2016}.
%Variational Autoencoders \parencite{kingma_auto-encoding_2013} are another popular method that is used for this sort of task. Their main drawback however, is that when a weak approximation of a prior or posterior distribution of the data is used, the model might learn something other than the true data distribution \parencite{goodfellow_nips_2016}.

The discovery of Generative Adversarial Networks \parencite[GANs;][]{goodfellow_generative_2014} has leapfrogged the state of developments in this field. Besides countering the drawbacks outlined above, they have been particularly successful at a variety of tasks such as image super-resolution \parencite[; see Figure \ref{fig:srgan}]{ledig_photo-realistic_2016}, style-based image generation \parencite{karras_style-based_2018}, or image-to-image translation \parencite[such as converting a satellite image into a map, or a sketch into a photorealistic image;][;see Figure \ref{fig:pix2pix}]{isola_image--image_2016}.

Similarly, and although not as popular, Recurrent Neural Networks (RNNs) have been successful at various generative tasks \parencite{jenal_rnn-based_2019, ICML2011Sutskever_524}. The reason is because RNNs contain a high-dimensional hidden state which allows them to remember and predict from previous inputs \parencite{ICML2011Sutskever_524}. This made them great candidates for predicting sequences \parencite{graves_generating_2013} of text or characters, where the input at $t_0$ might be influenced by prior input at $t_{-1}$, $t_{-2}$, etc. \parencite{fan_tts_2014}. That being said, Long Short-Term Memory networks \parencite[LSTM;][]{hochreiter_long_1997} - a variant of RNNs - have been purpose-built with memory cells which permit better storing and accessing of information. This design decision variant of the recurrent network good candidates for generating sequences with long-term dependencies \parencite{graves_generating_2013}.

\begin{figure}[t]
    \centering
    \begin{subfigure}{.24\linewidth}
        \includegraphics[width=\linewidth]{images/comic_HR.jpg}
        \caption{original}
        \label{fig:srgan_orig}
    \end{subfigure}
    \begin{subfigure}{.24\linewidth}
        \includegraphics[width=\linewidth]{images/comic_SRF_4_bicubic.jpg}
        \caption{bicubic}
        \label{fig:srgan_bicubic}
    \end{subfigure}
    \begin{subfigure}{.24\linewidth}
        \includegraphics[width=\linewidth]{images/comic_SRResNet-MSE.jpg}
        \caption{SRResNet}
        \label{fig:srgan_srresnet}
    \end{subfigure}
    \begin{subfigure}{.24\linewidth}
        \includegraphics[width=\linewidth]{images/comic_SRGAN-VGG54.jpg}
        \caption{SRGAN}
        \label{fig:srgan_srgan}
    \end{subfigure}

    \caption{Example of single-image super-resolution results that highlight the advantages of a GAN. The original high-resolution image (\ref{fig:srgan_orig}) has been downsampled to make a low-resolution image. The results of bicubic interpolation (\ref{fig:srgan_bicubic}) can be seen in the second image, followed by the results of SRResNet (\ref{fig:srgan_srresnet}), a neural network trained on mean squared error. Lastly, the results of SRGAN can be seen in the final image (\ref{fig:srgan_srgan}). Figure adapted from \textcite{ledig_photo-realistic_2016}.}
    \label{fig:srgan}
\end{figure}

\begin{figure}[t]
    \centering
    \begin{tabular}{ccc}
    Input & Ground Truth & Output \\
    \includegraphics[width=.15\linewidth]{images/handbags_edges_lotsofresults_latex/input_106_AB.jpg} &
    \includegraphics[width=.15\linewidth]{images/handbags_edges_lotsofresults_latex/gt_106_AB.jpg} &
    \includegraphics[width=.15\linewidth]{images/handbags_edges_lotsofresults_latex/L1cGAN_106_AB.jpg} \\
    \includegraphics[width=.15\linewidth]{images/handbags_edges_lotsofresults_latex/input_12_AB.jpg} &
    \includegraphics[width=.15\linewidth]{images/handbags_edges_lotsofresults_latex/gt_12_AB.jpg} &
    \includegraphics[width=.15\linewidth]{images/handbags_edges_lotsofresults_latex/L1cGAN_12_AB.jpg} \\
    \includegraphics[width=.15\linewidth]{images/handbags_edges_lotsofresults_latex/input_130_AB.jpg} &
    \includegraphics[width=.15\linewidth]{images/handbags_edges_lotsofresults_latex/gt_130_AB.jpg} &
    \includegraphics[width=.15\linewidth]{images/handbags_edges_lotsofresults_latex/L1cGAN_130_AB.jpg} \\
    \end{tabular}
    \caption{\textcite{isola_image--image_2016} coined the term image-to-image translation where an input image is transformed into a revisualization of the original input. This figures illustrates how an outline of a purse (leftmost column) is transformed into a colored photorealistic version of it (rightmost column). In the middle we have the ground truth: the real image of the purse. Figure adapted from \textcite{isola_image--image_2016}.}
    \label{fig:pix2pix}
\end{figure}

Having said all of these, the majority of developments with generative models lie in the field of Computer Vision. But that is not to say that we can generate only images. Generative models have been applied with varying degrees of success also in Natural Language Processing - where the scope was to generate high-quality text with sufficient diversity \parencite[e.g.,][]{yu_seqgan_2016, chen_adversarial_2018}. Less explored, but still successful, have been the efforts in generating music \parencite[e.g.,][]{mogren_c-rnn-gan_2016, dong_musegan_2017}. Herein lies the scope of this paper. Given such minimal explorations into generating music, we evaluated the samples generated by a GAN \parencite{mogren_c-rnn-gan_2016} and a LSTM network \parencite{oore_this_2018} trained under the constraints imposed by few-shot learning (Section \ref{sec:hungry}) against a baseline model proposed by \textcite{larochelle_few-shot_2017}. To reiterate, a full explanation of the purpose of the paper is provided in Section \ref{sec:thesis_goal}.

\section{I want it all}\label{sec:hungry}
% Queen - I want it all

DL models usually rely on a procedure called stochastic gradient descent (SGD) which implies computing an output and an error for a given input vector, followed by calculating the input's average gradient and adjusting the network's parameters in a manner that minimizes the error \parencite{lecun_deep_2015}. \textcite{ravi_optimization_2016} argue that the iterative nature of gradient-optimization algorithms does not allow them to perform well under the constraint of a set number of updates over few examples. In particular, when faced with non-convex optimization problems, DL algorithms do not guarantee convergence within a reasonable amount of time. In addition, they also emphasize that for each new problem (e.g., new dataset), the model needs to reinitialize with a new set of random parameters. Ultimately, this hurts the network's ability to converge to a good solution when constrained by a limited number of updates \parencite{ravi_optimization_2016}. Thus they lay out two clear drawbacks of DL models: "data-hunger", and inability to perform separate tasks.

A clear outcome from solving these issues would be that DL models are easier to train (less data) as well as multifunctional. Nonetheless, this list is not exhaustive. There is also fact that humans have the ability to generalize after one (or few) example(s) of a given object \parencite{vinyals_matching_2016, chen_closer_2018, ravi_optimization_2016} - something which DL models tend to lack. Moreover, there are many fields where the data exhibits a large number of classes, with few examples per class. Bridging the gap between human-type learning and current learning architectures would allow a system to do a proper job at capturing and generalizing information from sparse data \parencite{ravi_optimization_2016, larochelle_few-shot_2017}. This has motivated a series of explorations in the direction of "few-shot" learning (i.e., learning a class from a few examples) or more broadly, into the field of "meta-learning" \parencite[learning to learn in hopes of generalizing to new tasks;][]{chen_closer_2018, vinyals_matching_2016, zhang_metagan_2018}.

In spite of its proposed benefits, developments and experiments with few-shot learning are still scarce \parencite{larochelle_few-shot_2017}. Moreover, evaluations are largely concentrated on image data \parencite[see][]{lake_omniglot_2019, clouatre_figr_2019, vinyals_matching_2016, chen_closer_2018, ravi_optimization_2016}. Having said that, few-shot experiments with generative networks are even less frequent; \textcite{clouatre_figr_2019}, \textcite{dong_musegan_2017} and \textcite{zhang_metagan_2018} being some of the early (and few) explorers in this direction.

\section{Composing like a human} \label{sec:thesis_goal}

At the time of writing, generative models and meta-learning are two exciting areas of research the intersection of which could be quite promising. However, to the knowledge of the researcher, no experiments which combine the both of them have been conducted in the musical domain. Thus, this paper examined the extent to which generative models can create novel and qualitative samples of music under the constraints posed by few-shot learning. In other words, it sought to evaluate the degree to which generative models can create novel and high-fidelity samples when training data is scarce. To avoid any ambiguity, novel is used here as new (i.e., not in the training set), and qualitative as consistent with common scales and structurally diverse (combines chord structures with single-note sequences and is not restricted to only a few notes).

This research will evaluate the generated samples of two state-of-the-art generative models (C-RNN-GAN of \cite{mogren_c-rnn-gan_2016} and Performance-RNN of \cite{oore_this_2018}) that have been trained under a meta-learning framework, Reptile \parencite{nichol_first-order_2018}, with a limited number of songs. The performances of these two models were compared to the baseline proposed by \textcite{larochelle_few-shot_2017} and to recorded performances played by expert piano players at the International e-Piano Competition \parencite{university_of_minnesota_international_2019}.

In other words, this study sought to answer the following research questions:

@@

The following chapters will discuss more about generative models, meta-learning, and some of the models and algorithms involved in this paper (Chapter \ref{chap:theoretical_framework}), followed by a detailed description of the setup of the experiments (Chapter \ref{chap:methods}), and the results obtained from them (Chapter \ref{chap:results}). Finally, Chapter \ref{chap:discussion} closes with a discussion of the results and limitations of the experimental setup, whereas Chapter \ref{chap:conclusion} concludes the findings of this paper.

@@

This chapter takes a deep-dive into the fields of meta-learning and generative networks in order to highlight their current state of affairs. This will lay the groundwork for the experiments detailed in Chapter \ref{chap:methods}. Section \ref{sec:rnn} will present the principal network architecture of this paper, the recurrent neural network, and argument for its use as the foundation of a generative network. Section \ref{sec:generative} will introduce Generative Adversarial Networks, a training method build specifically for the task, whereas Section \ref{sec:learn2learn} will formally introduce meta-learning and some state-of-the-art approaches.

@@

For transparency, it should be noted that it is possible to use a feedforward architecture, specifically a one-dimensional Convolutional Neural Network, with sequential data, where the convolution forms the basis of a time-delay neural network. Unfortunately, the output of such a system would be a function of a small number of neighbouring data points \parencite[such as the last and the next three notes in a song;][]{goodfellow_deep_2016}. Despite that, a system that uses convolution would quickly reach its limitation in the musical domain. Here, knowledge of the notes played from the beginning of the song is needed, meaning that we need an architecture that is able to account for long-term dependencies - which brings us back to the RNN.

The information passes through the RNN in a similar manner to the feedforward network except that the activations which arrive at the hidden layer are from the current external input and hidden activations from the previous timestep. It should be noted that the parameters of a recurrent network (weights) are shared across different timesteps \parencite{graves_supervised_2012}. This allows extending (and applying) the model to data points of different lengths whilst generalizing across all of them \parencite{goodfellow_deep_2016}. To place this into perspective, a feedforward network would require separate parameters for each input feature, at each position.

The reason why these notions have been introduced is to be able to bring up the following point. As good as standard RNNs are - in theory - at accessing contextual information when mapping input sequences to output sequences, the range of contexts that can be accessed in a standard RNN architecture is in practice quite limited \parencite{graves_supervised_2012}. \textcite{hochreiter_long_1997} outlines two problems caused by BPTT: the error signals flowing backwards in time tend to either (1) blow up or (2) vanish. As such, the Long Short-Term Memory network has been proposed, a recurrent architecture designed to overcome these issues \parencite{hochreiter_long_1997}.

\begin{figure}[t]
    \centering
    \begin{tikzpicture}
        \node[io] (x1) {$x_1$};
        \node[io, above=4em of x1] (x2) {$x_2$};
        \node[io, above=4em of x2] (x3) {$x_3$};

@@

        \foreach \h in {1, 2, 3}
            \draw[conn] (h\h) -- (y);
    \end{tikzpicture}
    \caption{Example of a feedforward neural network with three nodes in the input layer ($x_1$, $x_2$, $x_3$), three nodes in the hidden layer ($h_1$, $h_2$, $h_3$), and an output layer a single node ($y$).}
    \label{fig:feedforward_net}
\end{figure}

\begin{figure}[t]
    \centering
    \begin{subfigure}{.24\linewidth}
        \begin{tikzpicture}
            [align=center]
            % Draw recurrent network
            \node[neuron] (h) {$h$};
            \node[io, below=4em of h] (x) {$x$};
            \node[io, above=4em of h] (y) {$y$};

@@

        \end{tikzpicture}
        \caption{Folded}
        \label{fig:rnn_folded}
    \end{subfigure}
    \begin{subfigure}{.74\linewidth}
        \begin{tikzpicture}
            [align=center]
            % Draw middle unfold
            \node[neuron] (ht) {$h^t$};
            \node[io, below=4em of ht] (xt) {$x^t$};
            \node[io, above=4em of ht] (yt) {$y^t$};

@@

        \end{tikzpicture}
        \caption{Unfolded}
        \label{fig:rnn_unfolded}
    \end{subfigure}
    \caption{A typical view of an RNN (left) and an unfolded view of the same network (right). In both sub-figures each node is a layer of network units however, in the second representation these are shown for each timestep. The weighted connections from the input layer to the hidden one is labelled $U$, those from the hidden layer feeding back to itself are $W$, and those from the hidden layer to the output layer are labelled as $V$. Note that these weights are reused at each timestep. Bias has been omitted in this representation. Figure adapted from \textcite{graves_supervised_2012}}
\end{figure}

@@

To make the generator's data distribution $p_g$ imitate that of the real data $p_{\text{data}}$, we need to define a prior on input noise variables $p_{z}(z)$ and then represent a mapping to the data space $G(z; \theta_g)$ where $G$ is a differentiable function represented by a network with parameters $\theta_g$. The discriminator is another network with a mapping function $D(x; \theta_d)$ that outputs a single scalar. More concretely, $D(x)$ is the probability that $x$ came from the real data distribution rather than that of the generator \parencite{goodfellow_nips_2016}. We train $D$ to maximize the probability of correctly classifying the training data and samples from $G$ (i.e., as "real" and "fake"), while simultaneously training $G$ to minimize the dissimilarity between the two data distributions, \emph{without actually looking at} $p_{\text{data}}$. In other words, $G$ must minimize $\log (1- D(G(z)))$ \parencite{goodfellow_generative_2014}.

@@

Although new, GANs are a significant conceptual and technical innovation \parencite{briot_deep_2017} which can be seen in them revitalizing interest in generative models. As mentioned before, they have been proven successful at image super-resolution \parencite{ledig_photo-realistic_2016}, image-to-image translation \parencite{isola_image--image_2016}, or style-based image generation \parencite{karras_style-based_2018}. Having said these, \textcite{mogren_c-rnn-gan_2016} has shown that GANs do not belong only to the field of Computer Vision by successfully re-purposing this architecture to generate music. The model in question, C-RNN-GAN, features a completely unidirectional LSTM network with two layers, 350 units each as generator and a unidirectional LSTM network with a similar structure as discriminator.

\section{Learning to learn} \label{sec:learn2learn}

The rationale for investing time and energy into meta-learning is that we cannot capture the essence of learning simply by relying on a small number of algorithms, but rather a bunch of context-dependent learning strategies that would enable us to properly capture domain specific information \parencite{schmidhuber_evolutionary_1987}. Given the complexities of each learning strategy, \textcite{schmidhuber_evolutionary_1987} argues that an obvious way to tackle this initiative is to give a system the ability to learn how to learn. A simplistic way to do this would be to train a system on different tasks (e.g, image classification, text classification, etc.). However, this approach entails training on a dataset that is much larger that what current hardware can withstand. Furthermore, this approach would stray away from the idea of building general-purpose A.I. architectures as generalizing over a task would require more data points that what a human would need. To be more concrete, a human is able to generalize a task/object after encountering one (or few) example(s) \parencite{ravi_optimization_2016, chen_closer_2018}. Therefore, besides having to perform on a variety of tasks, a system would also need to excel at that using sparse amounts of data.

Although ideas of meta-learning can go as far as Schmidhuber's thesis in 1987, it seems that training a model using a handful of examples did not get much traction until \textcite{lake_human-level_2015} successfully designed a probabilistic learning framework capable of generalizing and learning a large number of concepts from a single class example (i.e., one-shot learning). For clarity, such a problem is typically defined as a $n$-shot $k$-way problem, where $n$ stands for the number of examples per class, and $k$ is the number of classes to be used in the training process.

>>>>

\usepackage[colorlinks=true, citecolor=black, linkcolor=black, urlcolor=black]{hyperref}

% Essentials
\usepackage{tabularx}
\usepackage{supertabular}
\usepackage{graphicx}
\usepackage[font=small, labelfont=bf]{caption}
\usepackage{subcaption}
\usepackage{lscape}
\usepackage[ruled]{algorithm2e}

@@

\newenvironment{acknowledgements}%
    {\cleardoublepage\thispagestyle{empty}\null\vfill\begin{center}%
    \bfseries Acknowledgements\end{center}}%
    {\vfill\null}

@@

\cleardoublepage
\begin{center}
    \vspace*{\fill}
\begin{figure}[ht]
    \centering
    \def\svgwidth{\linewidth}
    \input{images/the_lick.pdf_tex}
\end{figure}
    \vspace{1em}
\end{center}

%TC:ignore
\chapter*{Abstract}
Experiments in Deep Learning are marked by scant explorations outside of the classification task, and even more so when trying to overcome some of the limitations imposed by gradient-based optimization algorithms (such as the uncertainty of convergence when dealing with non-convexity). With the resurgence of generative models and recent advancements in meta-learning we are faced two exciting ares of research, the intersection of which could be a promising area when investing general-purpose A.I. systems. Inspired by recent literature, this paper proposed a new line of inquiry where these two fields could be useful specifically, in the musical domain. As such, the paper sought to establish the extent to which meta-learned generative models are comparable to a model trained under normal circumstances, and to real musical performances.

@@

When compared to the baseline, the two Reptile-based models have shown to be quite competitive however, they failed to attain the same level of mastery when compared to the test samples. However, it is hypothesized that more time involved into training them could narrow the gap between the models and the real musical performances. Thus, the results of these experiments solidify the intuition that meta-learning algorithms are a viable alternative to traditional training methods, in the context of generative models.

@@

\chapter{Introduction} \label{chap:intro}

@@

Broadly speaking, DL as a field, promises to solve tasks which are easily (intuitively) performed by people, but hard to formalize \parencite[e.g., recognizing faces or spoken word;][]{goodfellow_deep_2016}. The solution comprises of enabling computers to learn from experience and understand the world through the discovery of hierarchical relationships between concepts \parencite{lecun_deep_2015}. This way, there is no need for manual input on the sort of knowledge that the computer needs \parencite{goodfellow_deep_2016}.

@@

\section{A variety of applications}\label{sec:no_disc}

@@

Regardless, successes in generative models have been either scarce or unattractive, with Fully Visible Belief Networks \parencite{frey_does_1996} being the most suitable (and popular) method for nearly two decades. Given that we have had a solution for all this time, a valid question to ask is why things haven't moved that much since then. As it turn out, this method requires generating one sample at a time (ramping up the cost of the process to $O(n)$, where $n$ is the number of samples). Coupled with the fact that the computation is done through a neural network, creating $n$ samples requires a substantial amount of time \parencite{goodfellow_nips_2016}. This discussion is not exhaustive; for a comprehensive list of various generative approaches and their downsides, refer to \textcite{goodfellow_nips_2016}.
%Variational Autoencoders \parencite{kingma_auto-encoding_2013} are another popular method that is used for this sort of task. Their main drawback however, is that when a weak approximation of a prior or posterior distribution of the data is used, the model might learn something other than the true data distribution \parencite{goodfellow_nips_2016}.

The discovery of Generative Adversarial Networks \parencite[GANs;][]{goodfellow_generative_2014} has leapfrogged the state of developments in this area. Besides countering the drawbacks outlined above, they have been particularly successful at a variety of tasks such as image super-resolution \parencite{ledig_photo-realistic_2016}, or style-based image generation \parencite{karras_style-based_2018}. Similarly, and although not as popular, Recurrent Neural Networks (RNNs) have been successful at various generative tasks \parencite{jenal_rnn-based_2019, ICML2011Sutskever_524}. The reason is because RNNs contain a high-dimensional hidden state which allows them to remember and predict from previous inputs \parencite{ICML2011Sutskever_524}. This made them great candidates for predicting sequences \parencite{graves_generating_2013} of text or characters, where the input at $t_0$ might be influenced by prior input at $t_{-1}$, $t_{-2}$, etc. \parencite{fan_tts_2014}. That being said, Long Short-Term Memory networks \parencite[LSTM;][]{hochreiter_long_1997} - a variant of RNNs - have been purpose-built with memory cells which permit better storing and accessing of information. This variant of the recurrent network makes it an excellent candidate for generating sequences with long-term dependencies \parencite{graves_generating_2013}.

In light of the above, the majority of developments with generative models lie in the field of Computer Vision. But that is not to say that we can generate only images. Generative models have also been applied with varying degrees of success also in Natural Language Processing - where the scope was to generate high-quality text with sufficient diversity \parencite[e.g.,][]{yu_seqgan_2016, chen_adversarial_2018}. Less explored, but still successful, have been the efforts in generating music \parencite[e.g.,][]{mogren_c-rnn-gan_2016, dong_musegan_2017}.

\textcite{briot_deep_2017} have conducted one of the most comprehensive survey of DL techniques and models for generating music, work which covers experiments from the last 25 years. What is striking, is that the authors admit that in spite of its rich history in computer science, there have been only a few attempts at generating music with ANNs. Thus, herein lies the scope of this paper. Given such minimal explorations, I evaluated the songs generated by a GAN \parencite{mogren_c-rnn-gan_2016} and a LSTM network \parencite{oore_this_2018} trained with a handful of examples against a baseline model that had access to the entire training set. To understand why such an assessment is needed, we need to introduce some of the shortcomings of modern ANNs and how to circumvent them.

\section{Pitfalls in gradient descent}\label{sec:hungry}

The training of DL models usually relies on a procedure called stochastic gradient descent (SGD) which implies computing an output and an error for a given input vector, followed by calculating the input's average gradient and adjusting the network's parameters in a manner that minimizes the error \parencite{lecun_deep_2015}. \textcite{ravi_optimization_2016} argue that the iterative nature of gradient-optimization algorithms does not allow them to perform well under the constraint of a set number of updates over few examples. In particular, when faced with non-convex optimization problems, DL algorithms do not guarantee convergence within a reasonable amount of time. In addition, they also emphasize that for each new problem (e.g., new dataset), the model needs to reinitialize with a new set of random parameters. Ultimately, this hurts the network's ability to converge to a good solution when constrained by a limited number of updates \parencite{ravi_optimization_2016}. Thus they lay out two clear drawbacks of DL models: "data-hunger", and performing separate tasks is problematic.

A clear outcome from solving these issues would be that DL models are easier to train (less data) as well as being able to carry out multiple tasks. But there is also the fact that humans have the ability to generalize after one (or few) example(s) of a given object \parencite{vinyals_matching_2016, chen_closer_2018, ravi_optimization_2016} - something which DL models tend to lack. Moreover, there are many fields where the data exhibits a large number of classes, with few examples per class. Bridging the gap between human-type learning and current learning architectures would allow a system to do a proper job at capturing and generalizing information from sparse data \parencite{ravi_optimization_2016, larochelle_few-shot_2017}. This has motivated a series of explorations in the direction of "few-shot" learning (i.e., learning a class from a few examples) or more broadly, into the field of "meta-learning" \parencite[learning to learn in hopes of generalizing to new tasks;][]{chen_closer_2018, vinyals_matching_2016, zhang_metagan_2018}.

In spite of its proposed benefits, developments and experiments with few-shot learning are still scarce \parencite{larochelle_few-shot_2017}. Moreover, evaluations are largely concentrated on image data \parencite[see][]{lake_omniglot_2019, clouatre_figr_2019, vinyals_matching_2016, chen_closer_2018, ravi_optimization_2016}, whereas those with music are virtually non-existent.

\section{Problem definition} \label{sec:thesis_goal}

At the time of writing, generative models and meta-learning are two exciting areas of research, the intersection of which could be interesting for A.I. research. Having said that, there are scarcely any few-shot experiments with generative networks; \textcite{clouatre_figr_2019}, \textcite{dong_musegan_2017} and \textcite{zhang_metagan_2018} being some of the early (and few) explorers in this direction. Moreover, to the knowledge of the researcher, no such experiments have been conducted in the musical domain to this day. Thus, to enrich the current literature of DL in music, this paper examines the extent to which generative models can create novel and qualitative samples of music under the constraints posed by few-shot learning. In other words, it seeks to evaluate the degree to which generative models can create novel and high-fidelity samples when trained in a manner which is (quantitatively) more akin to that of an aspiring musician. To avoid any ambiguity, novel is used here as new (i.e., not in the training set), and qualitative as consistent with common scales and structurally diverse (combines chord structures with single-note sequences and is not restricted to only a few notes).

This research evaluates the generated samples of two state-of-the-art generative models (C-RNN-GAN of \cite{mogren_c-rnn-gan_2016} and Performance-RNN of \cite{oore_this_2018}) that have been trained under a meta-learning framework, Reptile \parencite{nichol_first-order_2018}, with a limited number of songs. The performances of these two models were compared to the baseline proposed by \textcite{larochelle_few-shot_2017} and to recorded performances played by expert piano players at the International e-Piano Competition \parencite{university_of_minnesota_international_2019}.

In other words, this study seeks to answer the following research questions:

@@

The following chapters will discuss generative models, meta-learning, and some of the models and algorithms involved in this paper in more depth (Chapter \ref{chap:theoretical_framework}), followed by a detailed description of the experimental setup (Chapter \ref{chap:methods}), and the results obtained from them (Chapter \ref{chap:results}). Finally, Chapter \ref{chap:discussion} closes with a discussion of the results and limitations of the experimental setup, whereas Chapter \ref{chap:conclusion} concludes the findings of this paper.

@@

To highlight the current state of affairs, this chapter takes a deep-dive into the fields of meta-learning and generative networks. This will lay the groundwork for the experiments detailed in Chapter \ref{chap:methods}. Section \ref{sec:rnn} will present the principal network architecture of this paper, the recurrent neural network, and argument for its use as the foundation of a generative network. Section \ref{sec:generative} will introduce Generative Adversarial Networks, a training method build specifically for the task, whereas Section \ref{sec:learn2learn} will formally introduce meta-learning and recent state-of-the-art approaches.

@@

For transparency, it should be noted that it is possible to use a feedforward architecture with sequential data. This would take the form of a Convolutional Neural Network where the convolution forms the basis of a time-delay system. Unfortunately, the output of such a system would be a function of a small number of neighbouring data points \parencite[such as the last and the next three notes in a song;][]{goodfellow_deep_2016}. Despite that, a system that uses convolution would quickly reach its limitation. Using music (the topic of this paper) as an example, knowledge of the notes played from the beginning of the song is needed. This means that we need an architecture that is able to account for long-term dependencies - which brings us back to the RNN.

The information passes through the RNN in a similar manner to the feedforward network except that the activations which arrive at the hidden layer are from the current external input and hidden activations from the previous timestep. It should be noted that the parameters (weights) of a recurrent network are shared across different timesteps \parencite{graves_supervised_2012}. This allows extending (and applying) the model to data points of different lengths whilst generalizing across all of them \parencite{goodfellow_deep_2016}. To place this into perspective, a feedforward network would require separate parameters for each input feature, at each position.

\begin{figure}[htbp]
    \begin{subfigure}{\linewidth}
        \centering
        \begin{tikzpicture}
            \node[io] (x1) {$x_1$};
            \node[io, above=4em of x1] (x2) {$x_2$};
            \node[io, above=4em of x2] (x3) {$x_3$};

@@

            \foreach \h in {1, 2, 3}
                \draw[conn] (h\h) -- (y);
        \end{tikzpicture}
        \caption{Feedforward Network}
        \label{fig:feedforward_net}
        \vspace{1em}\break
    \end{subfigure}
    \begin{subfigure}{.24\linewidth}
        \begin{tikzpicture}
            [align=center]
            % Draw recurrent network
            \node[neuron] (h) {$h$};
            \node[io, below=4em of h] (x) {$x$};
            \node[io, above=4em of h] (y) {$y$};

@@

        \end{tikzpicture}
        \caption{Folded RNN}
        \label{fig:rnn_folded}
    \end{subfigure}
    \begin{subfigure}{.74\linewidth}
        \begin{tikzpicture}
            [align=center]
            % Draw middle unfold
            \node[neuron] (ht) {$h^t$};
            \node[io, below=4em of ht] (xt) {$x^t$};
            \node[io, above=4em of ht] (yt) {$y^t$};

@@

        \end{tikzpicture}
        \caption{Unfolded RNN}
        \label{fig:rnn_unfolded}
    \end{subfigure}

    \caption{Example of the two main ANN structures: a feedforward neural network (top), and a RNN (bottom). The feedforward neural network features three nodes in the input layer ($x_1$, $x_2$, $x_3$), three nodes in the hidden layer ($h_1$, $h_2$, $h_3$), and an output layer with a single node ($y$). On the bottom left we have a typical view of a RNN followed by an unfolded view of the same network (bottom right). In both sub-figures at the bottom each node is a layer of network units however, in the second representation these are shown for each timestep. The weighted connections from the input layer to the hidden one is labelled $U$, those from the hidden layer feeding back to itself are $W$, and those from the hidden layer to the output layer are labelled as $V$. Note that these weights are reused at each timestep. Bias has been omitted in this representation. RNN figure adapted from \textcite{graves_supervised_2012}}
    \label{fig:two_nets}
\end{figure}

The reason why these notions have been introduced is to be able to bring up the following point. As good as standard RNNs are - in theory - at accessing contextual information when mapping input sequences to output sequences, the range of contexts that can be accessed in a standard RNN architecture is in practice quite limited \parencite{graves_supervised_2012}. \textcite{hochreiter_long_1997} outline two problems caused by BPTT: the error signals flowing backwards in time tend to either (1) blow up or (2) vanish. As such, the Long Short-Term Memory network has been proposed, a recurrent architecture designed to overcome these issues \parencite{hochreiter_long_1997}.

@@

To make the generator's data distribution $p_g$ imitate that of the real data $p_{\text{data}}$, we need to define a prior on input noise variables $p_{z}(z)$ and then represent a mapping to the data space $G(z; \theta_g)$ where $G$ is a differentiable function represented by a network with parameters $\theta_g$. The discriminator is another network with a mapping function $D(x; \theta_d)$ that outputs a single scalar \parencite{goodfellow_generative_2014}. More concretely, $D(x)$ is the probability that $x$ came from the real data distribution rather than that of the generator \parencite{goodfellow_nips_2016}. We train $D$ to maximize the probability of correctly classifying the training data and samples from $G$ (i.e., as "real" and "fake"), while simultaneously training $G$ to minimize the dissimilarity between the two data distributions, \emph{without actually looking at} $p_{\text{data}}$. In other words, $G$ must minimize $\log (1- D(G(z)))$ \parencite{goodfellow_generative_2014}.

@@

Although new, GANs are a significant conceptual and technical innovation \parencite{briot_deep_2017} which can be seen in them revitalizing interest in generative models. They have been shown successful at image super-resolution \parencite{ledig_photo-realistic_2016}, image-to-image translation \parencite{isola_image--image_2016}, or style-based image generation \parencite{karras_style-based_2018}. Having said these, \textcite{mogren_c-rnn-gan_2016} has shown that GANs do not belong only to the field of Computer Vision by successfully re-purposing this architecture to generate music. The model in question, C-RNN-GAN, features a completely unidirectional LSTM network with two layers, 350 units each as generator and a bidirectional LSTM network with a similar structure as discriminator.

\section{Meta-Learning} \label{sec:learn2learn}

The rationale for investing time and energy into meta-learning is that we cannot capture the essence of learning simply by relying on a small number of algorithms, but rather a bunch of context-dependent learning strategies that would enable us to properly capture domain specific information \parencite{schmidhuber_evolutionary_1987}. Given the complexities of each learning strategy, \textcite{schmidhuber_evolutionary_1987} argues that an obvious way to tackle this initiative is to give a system the ability to learn how to learn. A simplistic way to do this would be to train a system on different tasks (e.g, image classification, text classification, etc.). However, this approach entails training on a dataset that is much larger than what current hardware can withstand. Furthermore, this approach would stray away from the idea of building general-purpose A.I. architectures as generalizing over a task would require more data points that what a human would need. To be more concrete, a human is able to generalize a task/object after encountering one (or few) example(s) \parencite{ravi_optimization_2016, chen_closer_2018}. Therefore, besides having to perform on a variety of tasks, a system would also need to excel at that using sparse amounts of data.

Although ideas of meta-learning can go as far as Schmidhuber's thesis in 1987, it seems that training a model using a handful of examples did not get much traction until \textcite{lake_human-level_2015} successfully designed a probabilistic learning framework capable of generalizing and learning a large number of concepts from a single class example (i.e., one-shot learning). For clarity, such a problem is typically defined as a $n$-shot $k$-way problem, where $n$ stands for the number of examples per class, and $k$ is the number of classes.
